Evaluating the Performance and Energy Efficiency of COSMO-ART, a Fully Online Coupled Model System with Numerical Weather Forecast and Chemical Transport Models

### 2014-04-01 trunk/paper_enahpc14/Sec2:Power-measurement.tex
adds missing files from previous commit

\subsection{Power-Performance measurement framework}

To assess the performance and the energy efficiency of COSMO-ART,
we employ a version of the integrated framework presented in~\cite{energy13}
that works in combination with VampirTrace and Vampir, which are profiling/tracing
and visualization tools, respectively. 
%The left part of the \vref{fig:Lustre} offers a graphical representation of
%the Lustre architecture; the right depicts the tracing and profiling framework.
To use our approach, COSMO-ART is compiled using the VampirTrace compiler wrappers, 
which automatically instrument the Fortran code of the model. Next, COSMO-ART 
is run on the nodes, thus dissipating certain amount of power.
The server nodes are connected to power measurement devices that account for
the dissipated power/consumed energy and send the power data to the tracing server.
The attached VampirTrace \pmlib plugin employs the client API that
sends start/stop primitives in order to gather captured data by the wattmeters onto
the tracing server, where an instance of the \pmlib server is running.
Once COSMO-ART run is finished, the VampirTrace \pmlib plugin
receives the power data from the tracing server.
The instrumentation post-process generates the performance trace files and the
\pmlib plugin inserts the power data into them.

In addition to the power measurements, we also account for the resource utilization
values of the nodes: CPU load, memory usage and storage device utilization. % and network utilization.
We run special \pmlib server instances on the server nodes
that retrieve these values from the \texttt{proc} file system (leveraging the \texttt{psutil} Python library).
Thus, \pmlib plugin instances running with the instrumented application
connect with the \pmlib servers. Finally, using the Vampir visualization tool, the power-performance traces
can be easily analyzed through a series of plots and statistics.

>>>>

\subsection{Power-Performance measurement framework}
\label{subsec:3.1}
To assess the  performance and the energy efficiency  of COSMO-ART, we
employ   a    version   of   the    integrated   framework   presented
in~\cite{energy13}  that  works in  combination  with VampirTrace  and
Vampir,   which  are   profiling/tracing   and  visualization   tools,
respectively.
%The left part of the \vref{fig:Lustre} offers a graphical representation of
%the Lustre architecture; the right depicts the tracing and profiling framework.
To  use our  approach,  COSMO-ART is  compiled  using the  VampirTrace
compiler wrappers, which automatically  instrument the Fortran code of
the  model. Next,  COSMO-ART is  run  on the  nodes, thus  dissipating
certain  amount of  power.  The  server nodes  are connected  to power
measurement  devices that  account for  the  dissipated power/consumed
energy and  send the power data  to the tracing  server.  The attached
VampirTrace \pmlib plugin employs the client API that sends start/stop
primitives in order to gather captured data by the wattmeters onto the
tracing server,  where an  instance of the  \pmlib server  is running.
Once COSMO-ART run is finished, the VampirTrace \pmlib plugin receives
the  power   data  from  the  tracing   server.   The  instrumentation
post-process  generates the  performance  trace files  and the  \pmlib
plugin inserts the power data into them.

In  addition  to the  power  measurements,  we  also account  for  the
resource utilization values  of the nodes: CPU load,  memory usage and
storage device utilization. % and network utilization.

We  run special  \pmlib  server  instances on  the  server nodes  that
retrieve these  values from the \texttt{proc}  file system (leveraging
the  \texttt{psutil} Python library).   Thus, \pmlib  plugin instances
running  with the  instrumented  application connect  with the  \pmlib
servers.   Finally,   using  the   Vampir   visualization  tool,   the
power-performance traces  can be easily  analyzed through a  series of
plots and statistics.
